A forest is more than its trees: haplotypes and inferred ARGs

### 2024-11-12 paper/writeup.tex
first final draft of results

% NOTE: these are in alphabetical order for now; we can revisit that ordering
\author[$\dagger$]{Halley Fritze}
\author[$\dagger$]{Nathaniel Pope}
\author[$\dagger$]{Ava Bamforth}
\author[$\ddagger$]{Jerome Kelleher}
\author[$\ast$,$\dagger$,S]{Peter Ralph}

@@

\runningtitle{Junctions in tree sequences}
\runningauthor{Bamforth \textit{et al.}}

%%%%%%%%%%
\begin{abstract}
    Foreshadowing haplotype-based methods of the genomics era,
    it is an old observation that the ``junction'' between two distinct haplotypes
    produced by recombination is inherited as a Mendelian marker.
    In this paper, we describe how this recombination-mediated information
    can in many cases be recovered from inference based solely on
    polymorphic markers (i.e., information produced by mutation),
    which produces unary nodes along the branches of the marginal genealogical trees.
    The resulting inferred ARGs, represented as tree sequences,
    are smaller, faster to compute with,
    and potentially contain substantially more information
    about the age of ancestral haplotypes,
    than inferred ARGs without these unary nodes.
    We provide efficient algorithms to identify this ancestral haplotype information,
    new metrics of agreement/disagreement between inferred ARGs,
    and explore some consequences for ARGs inferred from real data.
\end{abstract}

@@

Since the algorithm needs to take multiple passes over the tree sequence
in each direction,
an important practical question for this algorithm is:
how passes do we need to do?
The algorithm is monotone (spans of ancestral nodes only increase),
so it is guaranteed to require only a finite number of passes,
but it is also not hard to construct pathological cases that require an arbitrary number of passes.
Counting edges for every iteration suggest that only at most five iterations are needed before the algorithm terminates. Indeed, for even larger sequences \tabref{tab:edge-counts} shows that 99\% of all changes to an ARG occur after the first iteration with the algorithm completing after four iterations. 
%% Find counter example to global minimum.

@@

@@

\comment{TODO: put in benchmarking figures}

In the simple example in \figref{fig:conceptual},
extending haplotypes replaces three edges
($0 \to 3$ and $3 \to 4$ on the left tree, and $0 \to 4$ on the right tree)
by two edges ($0 \to 3$ and $3 \to 4$ on both trees).
If all edge endpoints were unique, then we'd expect \emph{every} edge to be extendable
on one of its ends
(except those pendant to the root and some of those adjacent to chromosome ends),
leading to a reduction in number of edges by 1/3.
Experiments with an earlier version of the algorithm showed that
if we only extend haplotypes on ``paths of length 1'' like this, 
then this is achieved for long sequences.
It is possible for \algorithmref{alg:extend} to add edges, as in \figref{fig:extending_diagram},
but we still expect the number of edges to decrease by more than 1/3.
Indeed,
\figref{fig:num_edges} shows that \algorithmref{alg:extend} nearly cuts the number of edges
in half, as long as the sequence is long enough.

This reduction in edges can also lead to a reduction in computation time.
Indeed, \figref{fig:num_edges} shows that computation time is reduced by 10--20\%
for a typical statistic (here, Tajima's $D$),
computed in an efficient iterative manner along the genome as implemented in \tskit.
This is actually somewhat surprising:
as described in \citet{ralph2020efficiently},
each edge added or removed requires an update
to the state of the parent node \emph{and every node above it in the tree}.
So, even if an edge extends for a long segment of genome,
something has to happen to it every time the subtree below it changes,
which is not affected by extending haplotypes.

@@

\comment{TODO: figure out if we can remove this from figures. If so, move this to Methods.}
Some stuff can't be inferred since it's "isolated unary";
for instance with 1,000 samples, 5e7 that's about half the ARG.
Etcetera.

@@

As seen in \figref{fig:node-spans},
Look, we add lots of stuff
and most most of it is right;
this applies to both recent and ancient nodes;
we're typically extending the span of a node by like maybe 50\%.

\begin{figure}[bht]
	\includegraphics[width=0.9\linewidth]{newplots_wo_ee/figure-4-2.pdf}
	\caption{
        Node spans of simulated tree sequence (blue), simplified tree sequence (green), and extended tree sequence (red).
        \comment{TODO: nodes should be in the same order in all three lines,
        using the ordering from the true tree sequence.}
        \comment{TODO: somehow communicate per-node amount/proportion of correctly and incorrectly added span}
    }
	\label{fig:node-spans}
\end{figure}

\begin{figure}
	\includegraphics{newplots_wo_ee/figure-4.pdf}
\end{figure}

\comment{merge this in}
The first thing we see is that coalescent nodes are unary
over a substantial portion of their spans:
since the \texttt{simplify} operations removes these unary spans,
the ``simplified'' line on the right of Figure~\ref{fig:discrepancy} shows
the proportion of nodes' spans on which they are not unary.
For instance, on an ARG of 10 samples on a sequence of length $5 \times 10^7$,
only about 15\% of the total spans are retained after simplification;
this rises to around 50\% for 1,000 samples (top right plot)

We can also use measures of dissimiliarity and similarity to quantify summaries across nodes:
these are shown in \figref{} (purple lines only)
Points:
proportion left after smiplify varies a LOT: from 15\% to 95\%.
    (goes up with L: with longer sequence length, greater opportunity for distant chunks to be on the same lineage)
extending adds hardly any wrong stuff (left plots); and adds a substantial chunk: with large samples, adds like half of the missing stuff

\begin{figure}
	\begin{center}
		% \includegraphics[width=0.8\linewidth]{newplots_wo_ee/discrepancy_over_sample_5e7_new.pdf}
		% \includegraphics[width=0.8\linewidth]{newplots_wo_ee/discrepancy_over_seqlen_1000s_new_5e7.pdf}
        \includegraphics[width=\textwidth]{newplots_wo_ee/dissimilarity.pdf}
	\end{center}
    \caption{
        Accuracy and sensitivity of extended haplotypes
        across a range of sample sizes, on a sequence of length $5\times10^7$ (top)
        and sequence lengths, with 1,000 \comment{diploid?} samples (bottom).
        For each, a simulated ARG containing unary haplotype spans
        was (i) \emph{simplified}, removing the unary spans,
        and (ii) \emph{inferred}, using \tsinfer{} on genotypes;
        then each of these had its haplotypes \emph{extended}.
        The inferred, then simplified, ARG is also shown.
        \textbf{Left:} Discrepancy to the true ARG,
        as proportion of haplotypes that are not represented in the true ARG
        (equation~\eqref{eq:discrepancy}).
        \textbf{Right:} Agreement to the true ARG,
        as proportion of the true ARG that is represented.
        The largest number of trees in the bottom row corresponds
        to the sequence length in the top row;
        see text for more details.
        \comment{TODO: make "inferred simplified" green (dotted) not red}
        \comment{TODO: label "inferred extended" "inferred simplfieid extended"??}
        \comment{TODO: also add a "inferred extended" line}
    }
    \label{fig:discrepancy}
\end{figure}

\begin{figure}
	\includegraphics{newplots_wo_ee/figure5.pdf}
\end{figure}

@@

So far, we have demonstrated that there is potentially ample information
in the coalescent-only trees to extend haplotypes.
Does this work with \emph{inferred} ARGs?
Indeed, most ARG inference methods provide inferred ARGS already containing unary nodes;
to what degree are they taking advantage of this information already?
Here, we study the behavior of \tsinfer{},
leaving a comparison of different ARG inference methods for future work.
To study this, we simulated ARGs containing unary-spanning haplotypes (as above),
then performed various operations to them,
reporting proportions of correct and incorrect haplotype spans in Figure~\ref{fig:discrepancy}.

Main points: in \figref{fig:discrepancy}
(green line) tsinfer gets a lot right (above 60\% with 1000 samples; bottom-right plot)
    and does a lot better with more samples;
    also, gets 10\%--25\% wrong (left panels)
    with longer sequence, larger proportion is wrong (bottom-left panel)
(red dotted)
    simplifying removes a lot of the wrong stuff
    but also a lot of the right stuff
(red solid)
    similarly extending the simplified one adds a bunch of right and also wrong stuff

@@

\begin{figure}
    \includegraphics{benchmarks/one_pop_results_ratios}
    \caption{
        Ratio of
        \textbf{(A)} number of edges, and
        \textbf{(B)} runtime for computing Tajima's $D$;
        before and after extending haplotypes.
        For instance, extending haplotypes reduces number of edges by about 50\%
        and statistic computation runtime by about 20\%
        for long sequences.
        Horizontal axis shows sequence length;
        colors show numbers of samples;
        with lines showing averages across replicates.
        The original tree sequence was simulated with an expanding population;
        see text for details. \comment{TODO}
        \label{fig:speed_and_edges}
    }
\end{figure}

@@

TODO

In this paper, we have taken a ``historical'' view of the ARG --
i.e., that each aspect of an inferred ARG is intended to represent
a portion of some particular historical genome
(for instance, the MRCA of some set of sampled genomes).
Furthermore, \figref{fig:conceptual}A\&B implicitly takes the position
that relationships \emph{not} depicted in the ARG are implied to not exist.
This is slighly disingenuous -- an alternative and more accurate interpetation
of the ARG of \figref{fig:conceptual}A\&B would be that we do not know
how node 2 inherited from node 4 on the right-hand interval,
rather than saying that it specifically did not pass through node 3.
The ``simplification'' algorithm of \citet{kelleher2018efficient},
or the Hudson algorithm for coalescent simulation
as implemented in msprime \citep{kelleher2016efficient},
each specifically discard information about any such ``non-coalescent'' nodes;
so the correct interpretation of these ARGs is a lack of knowledge.
In this paper, we have shown that this information can be largely replaced.

@@

\appendix

\begin{figure}
    \includegraphics{benchmarks/one_pop_results_absolute_values}
    \caption{
        \label{fig:speed_and_edges_absolute}
    }
\end{figure}

\begin{figure}
    \includegraphics{benchmarks/one_pop_results_timing}
    \caption{
        \label{fig:speed_and_edges_timing}
    }
\end{figure}

\begin{figure}
    \includegraphics{benchmarks/constant_pop_results_ratios}
    \caption{
        As in \figref{fig:speed_and_edges} except that
        the original tree sequence was simulated with an constant population;
        see text for details. \comment{TODO}
        \label{fig:speed_and_edges_constant}
    }
\end{figure}

\begin{figure}
    \includegraphics{benchmarks/constant_pop_results_absolute_values}
    \caption{
        \label{fig:speed_and_edges_absolute_constant}
    }
\end{figure}

\begin{figure}
    \includegraphics{benchmarks/constant_pop_results_timing}
    \caption{
        \label{fig:speed_and_edges_timing_constant}
    }
\end{figure}

>>>>

\author[$\dagger$]{Halley Fritze}
\author[$\dagger$]{Nathaniel Pope}
\author[$\ddagger$]{Jerome Kelleher}
\author[$\ast$,$\dagger$,S]{Peter Ralph}

@@

\runningtitle{Haplotypes in ARGs}
\runningauthor{Fritze \textit{et al.}}

%%%%%%%%%%
\begin{abstract}
    Foreshadowing haplotype-based methods of the genomics era,
    it is an old observation that the ``junction'' between two distinct haplotypes
    produced by recombination is inherited as a Mendelian marker.
    In this paper, we describe how this recombination-mediated information
    can in many cases be recovered from inference based solely on polymorphic markers,
    which produces unary nodes along the branches of the marginal genealogical trees.
    The resulting inferred ARGs are smaller, faster to compute with,
    and contain substantially more information
    than inferred ARGs without these unary nodes.
    We provide efficient algorithms to identify this ancestral haplotype information,
    new metrics of agreement and disagreement between inferred ARGs,
    and explore some consequences for ARGs inferred from real data.
\end{abstract}

@@

Since the algorithm needs to take multiple passes over the tree sequence
in each direction,
an important practical question for this algorithm is:
how passes do we need to do?
The algorithm is monotone (spans of ancestral nodes only increase),
so it is guaranteed to require only a finite number of passes,
but it is also not hard to construct pathological cases that require an arbitrary number of passes.
Counting edges for every iteration suggest that only at most five iterations are needed before the algorithm terminates.
Indeed, for even larger sequences \tabref{tab:edge-counts}
shows that 99\% of all changes to an ARG occur after the first iteration,
with the algorithm completing after four iterations.

@@

We used simulations of several scenarios.
To include the effects of heterogeneous recombination rate,
in some we used stdpopsim \citep{stdpopsim} 
to simulate chromosome 1 of \textit{Canis familiaris}
using the CanFam3 genetic map from \citet{campell2016}.
\emph{``Constant dog''} simulations simulated this chromosome in a 
population of (constant) size $10^4$.
\emph{``Expanding dog''} simulations were similar, but
used a discrete-time Wright--Fisher model to simulate
a small population of 100 that expanded to 1,000 individuals ten generations ago,
which then doubled every generation to reach 512,000 individuals.
All jobs used for recording runtime were run on an Intel Xeon Gold 6148 processor.
\comment{TODO: describe sims for inference?}

@@

In the simple example in \figref{fig:conceptual},
extending haplotypes replaces three edges
($0 \to 3$ and $3 \to 4$ on the left tree, and $0 \to 4$ on the right tree)
by two edges ($0 \to 3$ and $3 \to 4$ on both trees).
If all edge endpoints were unique, then we'd expect \emph{every} edge to be extendable
on one of its ends
(except those pendant to the root and some of those adjacent to chromosome ends),
leading to a reduction in number of edges by 1/3.
Experiments with an earlier version of the algorithm showed that
if we only extend haplotypes on ``paths of length 1'' like this, 
then this is achieved for long sequences.
It is possible for \algorithmref{alg:extend} to add edges, as in \figref{fig:extending_diagram},
but we still expect the number of edges to decrease by more than 1/3.
Indeed,
\figref{fig:speed_and_edges} shows that \algorithmref{alg:extend} nearly cuts the number of edges
in half, as long as the sequence is long enough.

\begin{figure}
    \centering
    \includegraphics{benchmarks/one_pop_results_ratios}
    \caption{
        Ratio of
        \textbf{(A)} number of edges, and
        \textbf{(B)} runtime for computing Tajima's $D$;
        before and after extending haplotypes.
        For instance, extending haplotypes reduces number of edges by about 50\%
        and statistic computation runtime by about 20\%
        for long sequences.
        Horizontal axis shows sequence length;
        colors show numbers of samples;
        with lines showing averages across replicates.
        The original tree sequence was simulated with the ``expanding dog'' expanding population
        and subset to various sizes;
        see Methods for details. 
        Absolute values are shown in Supplementary \figref{fig:speed_and_edges_other}.
        \label{fig:speed_and_edges}
    }
\end{figure}

This reduction in edges can also lead to a reduction in computation time.
Indeed, \figref{fig:speed_and_edges} shows that computation time is reduced by 10--20\%
for a typical statistic (here, Tajima's $D$),
computed in an efficient iterative manner along the genome as implemented in \tskit.
This is actually somewhat surprising:
as described in \citet{ralph2020efficiently},
each edge added or removed requires an update
to the state of the parent node \emph{and every node above it in the tree}.
So, even if an edge extends for a long segment of genome,
something has to happen to it every time the subtree below it changes,
which is not affected by extending haplotypes.

Supplementary \figref{fig:speed_and_edges_constant} shows these results are not specific
to the demographic scenario.
Supplementary Figures~\ref{fig:speed_and_edges_other}
and~\ref{fig:speed_and_edges_other_constant}
also show that our implementation of \algorithmref{alg:extend} is quite efficient,
running at chromosome scale in seconds to minutes for hundreds or thousands of samples,
or minutes to hours for tens of thousands of samples

@@

@@

As seen in \figref{fig:node-spans},
the vast majority of span added by extending haplotypes is correct.
In this example (which is typical),
99\% of all added span is correct;
95\% of nodes have no incorrectly added span;
and those incorrectly added spans are nearly always a small fraction of the original span.
The added information is significant:
the algorithm typically increases spans (i.e., lengths of ancestral haplotypes)
by around 50\%.

\begin{figure}[bht]
	\includegraphics{newplots_wo_ee/figure-4.pdf}
	\caption{
        The effect of extending haplotypes on per-node spans
        in an ARG simulated \comment{TODO HOW}.
        \textbf{(A)}
        Distribution of total amount of span added across nodes by \algorithmref{alg:extend};
        note the log scale on the $y$ axis.
        \textbf{(B)}
        Amount of incorrectly added span, plotted against total span, by node.
        95\% of nodes have no incorrect span; of the remainder,
        nearly all have less than 5\% incorrectly added;
        see Supplementary \figref{fig:incorrect_ratio}.
        Plots \textbf{(C)} and \textbf{(D)} show total spans per node,
        ordered by total span in the original ARG (which includes unary nodes).
        Dark blue dots in both show the spans in this original ARG.
        In \textbf{(C)}, lighter green dots show
        span after removing unary spans using \textit{simplify},
        while in \textbf{(D)},
        lighter red dots show span after extending haplotypes,
        i.e., applying \algorithmref{alg:extend} to the simplified ARG.
    }
	\label{fig:node-spans}
\end{figure}

These statistics are also reflected at the genomic scale, using measures of dissimilarity.
\figref{fig:dissimilarity} shows in purple total amounts of span removed by simplification
and re-inferred by \algorithmref{alg:extend} (correctly and incorrectly).
(Lines in green involve re-inference of the ARG; discussed next.)
The top row shows the proportion of the given ARG that does not match the original,
showing that the total amount of mis-matching span
produced by extending haplotypes is very small ($\approx 1\%$).
(Simplification does not produce dissimilarity.)
The bottom row shows the proprtional of the original ARG that is represented in the given ARG.
This shows that a large proportion of the ARG can be removed by simplifiction,
indicating that coalescent nodes are unary
over a substantial portion of their spans.
(Since the \texttt{simplify} operations removes these unary portions of haplotypes,
the ``simplified'' line (``SS'') on the right of Figure~\ref{fig:dissimilarity} shows
the proportion of nodes' spans on which they are not unary.)
However, \algorithmref{alg:extend} can correctly replace most of these portions of haplotypes,
especially with larger sample sizes and sequence lengths
(the ``simplified--extended'' line; ``SSE'').
For instance, with 1,000 samples and a $5 \times 10^7$bp genome,
haplotypes are unary on about half their spans (on average),
and extending haplotypes can infer more than half of this span from coalescent information only.

\begin{figure}
	\begin{center}
		% \includegraphics[width=0.8\linewidth]{newplots_wo_ee/discrepancy_over_sample_5e7_new.pdf}
		% \includegraphics[width=0.8\linewidth]{newplots_wo_ee/discrepancy_over_seqlen_1000s_new_5e7.pdf}
        \includegraphics[width=\textwidth]{newplots_wo_ee/dissimilarity.pdf}
	\end{center}
    \caption{
        Accuracy and sensitivity of extended haplotypes
        across a range of sample sizes, on a sequence of length $5\times10^7$ (top)
        and sequence lengths, with 1,000 \comment{diploid?} samples (bottom).
        For each, a simulated ARG containing unary haplotype spans
        was (i) \emph{simplified}, removing the unary spans,
        and (ii) \emph{inferred}, using \tsinfer{} on genotypes;
        then each of these had its haplotypes \emph{extended}.
        The inferred, then simplified, ARG is also shown.
        \textbf{Left:} Discrepancy to the true ARG,
        as proportion of haplotypes that are not represented in the true ARG
        (equation~\eqref{eq:discrepancy}).
        \textbf{Right:} Agreement to the true ARG,
        as proportion of the true ARG that is represented.
        The largest number of trees in the bottom row corresponds
        to the sequence length in the top row;
        see text for more details.
        \comment{TODO: make "inferred simplified" green (dotted) not red}
        \comment{TODO: also add a "inferred extended" line}
        \comment{TODO: instead of `SS` have `S` ; and instead of `SSE` have `SE`.}
    }
    \label{fig:dissimilarity}
\end{figure}

\begin{figure}
	\includegraphics{newplots_wo_ee/figure5.pdf}
    \caption{I think this is being removed?}
\end{figure}

@@

So far, we have demonstrated that there is potentially ample information
in the coalescent-only trees to extend haplotypes.
Does this work with \emph{inferred} ARGs?
Indeed, most ARG inference methods provide inferred ARGS already containing unary nodes;
to what degree are they taking advantage of this information already?
Here, we study the behavior of \tsinfer{},
leaving a comparison of different ARG inference methods for future work.
To study this, we simulated ARGs containing unary-spanning haplotypes (as above),
then re-inferred ARGs from their genotypes and performed various operations on the results.

First, \figref{fig:dissimilarity} shows that \tsinfer{}
has a substantial portion of already--extended haplotypes:
comparing the ``inferred'' (``I''; dotted) line to the ``inferred-simplified'' (``IS''; dashed) line
we see that inclusion of these unary spans increases the amount of correctly inferred material
by around 20\% (bottom panels),
but that roughly 10\% of the unary spans in the inferred ARG are incorrect (top panels).
Applying \algorithmref{alg:extend} to the inferred-and-then-simplified ARG
(``ISE''; solid green line)
produces an ARG with both less correct and incorrect span.
It is also helpful to note that accuracy
(i.e., proportion of the true ARG that is inferred; bottom panels)
greatly increases with larger sample sizes,
possibly due to resolution of polytomies.
(Recall that due to computational constraints,
``correct'' and ``incorrect'' spans are determined here by \emph{exact} match
of subtending samples.)

reporting proportions of correct and incorrect haplotype spans in Figure~\ref{fig:dissimilarity}.
Main points: in \figref{fig:dissimilarity}
(green line) tsinfer gets a lot right (above 60\% with 1000 samples; bottom-right plot)
    and does a lot better with more samples;
    also, gets 10\%--25\% wrong (left panels)
    with longer sequence, larger proportion is wrong (bottom-left panel)
(red dotted)
    simplifying removes a lot of the wrong stuff
    but also a lot of the right stuff
(red solid)
    similarly extending the simplified one adds a bunch of right and also wrong stuff

@@

@@

We began work on this paper with the simple observation that the simple transformation
of \figref{fig:extending_diagram} would reduce the number of edges
in the succinct tree sequence representation of the ARG.
This is essentially a recombination-based parsimony argument,
and we have shown that this line of reasoning
leads to ARGs that are substantially more compact and faster to operate on,
and that contain more accurate information about true ancestral relationships.
These extended ancestral haplotypes manifest as unary nodes in the marginal trees.
Although a number of ARG inference methods may be taking advantage of this information,
it is our impression that this source of information is not widely appreciated in the community.
In fact, the field's focus on marginal trees rather than haplotypes
meant that to study accuracy of this method,
we had to invent new haplotype-aware measures of (dis)agreement between ARGs.

There are good reasons to think that lengthening the spans of ancestral haplotypes
could lead to substantial gains in accuracy of ARG inference.
For instance, information about inferring the age of a particular mutation
derives almost entirely from constraints at nearby, linked sites.
Extending ancestral haplotypes from one site into neighboring regions
conceptually allows information from those marginal trees
to inform age inference at that site as well.

We have also explored the degree to which \tsinfer{} already makes use of this information,
and whether this algorithm can be used to improve inference.
The results do not provide a clear ordering:
for instance, although \tsinfer-produced ARGs
have a substantial portion of correctly inferred unary haplotypes,
removing this with simplification decreases both dissimilarity
(i.e., proportion of ``wrong'' haplotypes)
and the proportion of the truth that is correctly inferred.
Extending haplotypes restores a large amount of this correctly inferred span,
but also introduces incorrect spans.
Further work is needed to determine how the balance of ``true and false positive rates''
affects downstream uses.
The efficient computational tools we have implemented (in \tskit{})
should facilitate this exploration.

\paragraph{Ignorance and omission in the ARG}
As motivation, we presented above a ``historical'' view of the ARG --
i.e., that each aspect of an inferred ARG is intended to represent
a portion of some particular historical genome
(for instance, the MRCA of some set of sampled genomes).
Furthermore, \figref{fig:conceptual}A\&B implicitly takes the position
that relationships \emph{not} depicted in the ARG are implied to not exist.
An alternative interpetation
of the ARG of \figref{fig:conceptual}A\&B would be that we do not know
how node 2 inherited from node 4 on the right-hand interval,
rather than saying that it specifically did not pass through node 3.
The ``simplification'' algorithm of \citet{kelleher2018efficient},
or the Hudson algorithm for coalescent simulation
as implemented in msprime \citep{kelleher2016efficient},
each specifically discard information about any such ``non-coalescent'' nodes;
so the correct interpretation of these ARGs is a lack of knowledge.
In this paper, we have shown that this information can be largely replaced.

@@

\appendix
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\begin{figure}
    \includegraphics{benchmarks/one_pop_results_absolute_values}
    \includegraphics{benchmarks/one_pop_results_timing}
    \caption{
        \textbf{(Above:)}
        Absolute values for ratios shown in \figref{fig:speed_and_edges}:
        \textbf{(left)} numbers of edges; and
        \textbf{(right)} runtime.
        \textbf{(Below:)} runtime for the \texttt{extend\_haplotypes}
        implementation of \algorithmref{alg:extend} provided in the \tskit{} library.
        For other details, see \figref{fig:speed_and_edges}.
        \label{fig:speed_and_edges_other}
    }
\end{figure}

\begin{figure}
    \includegraphics{benchmarks/constant_pop_results_ratios}
    \caption{
        As in \figref{fig:speed_and_edges} except that
        the original tree sequence was simulated with an constant population
        of size $10^4$ (using the ``constant dog'' scenario);
        see Methods for details.
        Absolute values are shown in Supplementary \figref{fig:speed_and_edges_constant_other}.
        \label{fig:speed_and_edges_constant}
    }
\end{figure}

\begin{figure}
    \includegraphics{benchmarks/constant_pop_results_absolute_values}
    \includegraphics{benchmarks/constant_pop_results_timing}
    \caption{
        As in \figref{fig:speed_and_edges_other} except that
        the original tree sequence was simulated with an constant population
        of size $10^4$ (using the ``constant dog'' scenario).
        \label{fig:speed_and_edges_other_constant}
    }
\end{figure}
